---
title: "EDA of Beer Production"
author: "Juan Solorio"
date: "4/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, load-libraries}
library(tidyverse)
require(forecast) #seasonplot
require(fda) #create bases for functional data
require(rainbow) #for coloring curves
require(qcc) #ewma control chart
require(car) #better QQ plots
data <- read.csv("../../Data/Beer.csv")
```
```{r, times-series}
data$DATE <- as.Date(data$DATE, format = "%m/%d/%Y")
beer.ts <- ts(data = data$BeerProduction, frequency = 12,start = c(1992,1),end = c(2015,12))
time.ts <- seq(as.Date("1992-01-01"), as.Date("2015-12-01"), by='1 month' )

# data frame with basic structures
ts.df = data.frame(Time = time.ts, Beer.Production = beer.ts)
ts.df$Year = as.integer(strftime(time.ts,'%Y'))
ts.df$Month = as.integer(strftime(time.ts,'%m'))
```

The data loaded is a data frame and needs to be converted into a 'time-series' object for us to perform some analysis.
Hense the `r ts` function takes in the `r BeerProduction` variable and makes this transformation.

```{r, plot1}
# autoplot takes in the time series object and plots it
autoplot(beer.ts) +
  ggtitle("Beer Production Time Series") +
  ylab("Beer Production") + xlab("Date")
```

From the chart we can see there is both a trend and seasonality.

```{r}
seasonplot(beer.ts)
```

```{r}
#look at annual data to better see smooth trend
beer_yr = aggregate(data = ts.df, Beer.Production ~ Year, sum  )
beer_mn = aggregate(data = ts.df, Beer.Production ~ Month, mean  )
ts.plot(beer_mn$Beer.Production)
ts.plot(beer_yr$Beer.Production)
vb_yr = sort(unique(beer_yr$Year))

```


```{r}
#check if monthly proportions roughly look constant year-to-year
colnames(beer_yr)[2] = "beer_annual"
ts.df = merge(ts.df, beer_yr, by = "Year")
ts.df$y_prop = ts.df$Beer.Production/ts.df$beer_annual

ts.plot(ts.df$y_prop)

```

Monthly propotion of the production of beer seem to be faily constant year-to-year according to the figure.

```{r}
#Look at distribution in monthly proportions (assuming stationary)
boxplot(data = ts.df, y_prop ~ Month) ##check similarity in proportions

```

```{r}
##### STL Decompostion
#given seasonal effect depends on level do log transform
beer_log = log(beer.ts)

#decompose and plot
fit_stl_beer = stl(beer_log, s.window = 12)
plot(fit_stl_beer)

#extract components
S_b = fit_stl_beer$time.series[,1]
L_b = fit_stl_beer$time.series[,2]
E_b = fit_stl_beer$time.series[,3]

ts.plot(exp(S_b), main = 'Multiplicative Seasonality Component')

```

```{r}
##### Functional Data View of Seasonality
#convert univariate series to 12 curves for each year
Mb = matrix(NA, nrow = 12, ncol =  length(vb_yr) )
ts.df$S = exp(S_b)

for( j in 1:ncol(Mb)  ){
  #subset
  d_j = subset(ts.df, Year == vb_yr[j])
  S_j = d_j$S
  S_j = S_j/mean(S_j) #ensure normalized
  
  #add
  Mb[,j] = S_j
}

colnames(Mb) = vb_yr

#Add Fourier Bases and plot
vb_f <- create.fourier.basis(  rangeval = c(0,nrow(Mb)),  nbasis = 7)
Sb_basis <- smooth.basis( y = Mb, fdParobj = vb_f)
plot(Sb_basis$fd)


#Add mean
Wb.mean <- mean.fd(Sb_basis$fd)
lines(Wb.mean, lty = 2, col='black', lwd=3)

```

```{r}
##### Look at Multiplicative Residuals
#plot centered about zero to eyeball randomness
ts.plot(E_b)
tbSeq = 1:length(E_b)
abline(h=0, col='black')

#autocorrelation
acf(E_b)

#control chart at slow smoothing
ewma(E_b, lambda = .2)

```

```{r}
#fit ar(1) model and extract residuals
fitb_ar1 = arima(E_b, order = c(1,0,0))
rb = fitb_ar1$residuals
rb = rb/sd(rb)

#distribution
hist(rb, main='Frequency of Normalized Multiplicative Residuals')

#compare fit against standard Gaussian
qqPlot(rb, main='Normalized Multiplicative Residuals vs N(0,1)')

##notice that variance does not constant. 

```
